---
output: 
  # bookdown::word_document2: 
  #   keep_tex: yes
  #   toc: false
  #   number_sections: false
  #   extra_dependencies: ["float"]
  #   always_allow_html: true
  bookdown::html_document2:
    toc: false
    fig_caption: yes
    number_sections: false
    keep_md: true
    self_contained: true
#bibliography: ecol.bib
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(emmeans)
library(here)
library(readr)
library(webshot2)
#need relative path for figures
knitr::opts_chunk$set(message = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                       webshot = TRUE,
                      fig.pos = "H", 
                      table.placement = "H",
                    #  out.extra = "",
                      dev=c("png")) 
                     # fig.width = 8, #change number specs to balance plot size and resolution
                    #  fig.height = 8,
                    #  fig.fullwidth = TRUE,
                     # fig.path = here("4 Graphs/"), 
                   #   dev=c("png"),
                   #   fig.pos = "H", 
                   #   table.placement = "H")
                   #   out.extra = "")

library(tidyverse)
library(magrittr)
library(lme4)
library(patchwork) # for wrap_plots
library(nlme)
library(kableExtra)
library(ggResidpanel)
library(gridExtra)
library(lemon)
library(data.table)
library(ggpubr)
theme_set(theme_bw())
```
 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
corn <- read_csv("../2-Data/Clean/M.corn_1720_clean.csv")
corn  %<>%
  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),
            funs(factor(.)))

soy <-  read_csv("../2-Data/Clean/M.soybean_1720_clean.csv")
soy  %<>% 
  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),
            funs(factor(.)))

oat <- read_csv("../2-Data/Clean/M.oat_g_1720_clean.csv")
oat  %<>%  
  mutate_at(c("Block","Crop_ID", "Year"),
            funs(factor(.)))

alfalfa <- read_csv("../2-Data/Clean/M.alfalfa_1720_clean.csv")
alfalfa %<>%  
  mutate_at(c("Block", "Year"),
                        funs(factor(.)))
```


```{r}
names(corn)
names(soy)
names(oat)
names(alfalfa)
```

### Yield models, bu/ac for grains
#### Corn

```{r}
table(corn$Crop_ID, corn$Year)
```

```{r, include=FALSE}
corn.lmer.bu <- lmer(`bu/acre at 15.5% moisture` ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)



corn.lmer.bu.ln <- lmer(log(`bu/acre at 15.5% moisture`) ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)

corn.lmer.bu.s <- lmer(sqrt(`bu/acre at 15.5% moisture`) ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)
```

#### Corn diagnosis: `corn.lmer` and `corn.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden corn yield. Model fitting does not differ between original vs. transform."}
list(corn.lmer.bu, corn.lmer.bu.ln, corn.lmer.bu.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

#### Soy

```{r}
table(soy$Crop_ID, soy$Year)
```

```{r}
soy.lmer.bu <- lmer(`bu/acre at 13% moisture` ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)

soy.lmer.bu.ln <- lmer(log(`bu/acre at 13% moisture`) ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)

soy.lmer.bu.s <- lmer(sqrt(`bu/acre at 13% moisture`) ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)
```


#### Soybean diagnosis: `soy.lmer` and `soy.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden soybean yield. Model fitting does not differ between original vs. transform."}
list(soy.lmer.bu, soy.lmer.bu.ln, soy.lmer.bu.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

#### Oat

```{r}
table(oat$Crop_ID, oat$Year)
```

```{r}
oat.lmer.bu <- lmer(`bu/ac at 14% moisture` ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)


oat.lmer.bu.ln <- lmer(log(`bu/ac at 14% moisture`) ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)


oat.lmer.bu.s <- lmer(sqrt(`bu/ac at 14% moisture`) ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)
```


#### Oat diagnosis: `oat.lmer` and `oat.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden oat yield. Model fitting does not differ between original vs. transform."}
list(oat.lmer.bu, oat.lmer.bu.ln, oat.lmer.bu.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```


### Yield models, ton/ac 

#### Corn

```{r}
table(corn$Crop_ID, corn$Year)
```

```{r, include=FALSE}
corn.lmer <- lmer(`ton/ac at 15.5% moisture` ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)



corn.lmer.ln <- lmer(log(`ton/ac at 15.5% moisture`) ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)

corn.lmer.s <- lmer(sqrt(`ton/ac at 15.5% moisture`) ~ Block +
                    Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=corn)
```

#### Corn diagnosis: `corn.lmer` and `corn.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden corn yield. Model fitting does not differ between original vs. transform."}
list(corn.lmer, corn.lmer.ln, corn.lmer.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

#### Soy

```{r}
table(soy$Crop_ID, soy$Year)
```

```{r}
soy.lmer <- lmer(`ton/ac at 13% moisture` ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)

soy.lmer.ln <- lmer(log(`ton/ac at 13% moisture`) ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)

soy.lmer.s <- lmer(sqrt(`ton/ac at 13% moisture`) ~ Block +  Crop_ID*Corn_weed_management + 
                    (1|Year) + 
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) +
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID),
  data=soy)
```


#### Soybean diagnosis: `soy.lmer` and `soy.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden soybean yield. Model fitting does not differ between original vs. transform."}
list(soy.lmer, soy.lmer.ln, soy.lmer.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

#### Oat

```{r}
table(oat$Crop_ID, oat$Year)
```

```{r}
oat.lmer <- lmer(`ton/ac at 14% moisture` ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)


oat.lmer.ln <- lmer(log(`ton/ac at 14% moisture`) ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)


oat.lmer.s <- lmer(sqrt(`ton/ac at 14% moisture`) ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID),
  data=oat)
```


#### Oat diagnosis: `oat.lmer` and `oat.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden oat yield. Model fitting does not differ between original vs. transform."}
list(oat.lmer, oat.lmer.ln, oat.lmer.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

#### Alfalfa

```{r}
table(alfalfa$Crop_ID, alfalfa$Year)
```

```{r}
## summing all the cuts within one year 
alf_sum <- alfalfa %>%
  group_by(Year, Block, Crop_ID) %>%
  summarize(Yield.ton.ac = sum(`ton/ac at 15% moisture`)) 
```


```{r}
table(alf_sum$Crop_ID, alf_sum$Year)
```

```{r}
alf.lmer <- lmer(Yield.ton.ac ~ Block + (1|Year),
  data=alf_sum)

alf.lmer.ln <- lmer(log(Yield.ton.ac) ~ Block + (1|Year),
  data=alf_sum)

alf.lmer.s <- lmer(sqrt(Yield.ton.ac) ~ Block + (1|Year),
  data=alf_sum)
```

#### Alfalfa diagnosis: `alf.lmer` and `alf.lmer.ln` similar

```{r, fig.height= 4, fig.width=12, fig.cap= "Model diagnosis for Marsden alfalfa hay yield. Model fitting does not differ between original vs. transform."}
list(alf.lmer, alf.lmer.ln, alf.lmer.s) %>%
  lapply(resid_panel,  qqbands = TRUE , nrow = 2) %>%
                              wrap_plots() +
                              plot_annotation(tag_levels = "a")
```

```{r , include=FALSE}
## ANOVA tables for the effects of crop identity and corn weed management on crop yields

corn_jt <- print(joint_tests(emmeans(corn.lmer.ln,
                                     c("Crop_ID", "Corn_weed_management"))), 
                 export = TRUE)
soy_jt <- print(joint_tests(emmeans(soy.lmer.ln, 
                                    c("Crop_ID", "Corn_weed_management"))),
                export = TRUE)

oat_jt <- print(joint_tests(emmeans(oat.lmer.ln, c("Crop_ID"))),export = TRUE)

#alf_jt <- print(joint_tests(emmeans(alf.lmer, c("Year"))),export = TRUE)



crop_jt <- rbind(corn_jt$summary,
                 soy_jt$summary,
                 oat_jt$summary)
  
rownames(crop_jt) <- NULL



#colnames(crop_jt) <- c("Model term", "df1", "df2","F","p")

## Merge 3 ANOVA tables  
crop_df <- as.data.frame(crop_jt)
#make sure that the F.value is numeric

crop_df$F.ratio <- as.numeric(crop_df$F.ratio)
```


### How did rotation system and corn weed management affect crop yields? {-}   

Results of the experiment indicated that crop diversification and reduced use of herbicides were not associated with lower crop yields (Table \@ref(tab:crop-jt-ct)). Averaged over four years, soybean was the only crop whose yield was affected by rotation (p = 0.0191, Table \@ref(tab:crop-jt-ct)). Soybean yield was 16% higher in the 4-year rotation than in the 2-year rotation (p = 0.0181). Crop yields in the experiment were as high or higher than the averages for the state of Iowa and Boone County (Figure \@ref(fig:crop-bar)). 

```{r crop-jt-ct, echo=FALSE}

## Rotation effect on corn yield

corn_ct <- print(contrast(emmeans(corn.lmer.ln, ~ Crop_ID, 
                                  infer = c(FALSE,TRUE), 
                                  type = "response"),
                          method = "pairwise"), 
                 export = TRUE)

## Rotation effect on soybean yield
soy_ct <- print(contrast(emmeans(soy.lmer.ln, ~ Crop_ID,
                                 infer = c(FALSE,TRUE),
                                 type = "response"),
                         method = "pairwise"),
                export = TRUE)


## Rotation effect on oat yield

oat_ct <- print(contrast(emmeans(oat.lmer.ln, ~ Crop_ID,
                                 infer = c(FALSE,TRUE), 
                                 type = "response"), 
                         method = "pairwise", 
                         adjust = "none"), #no adjustment to get p-value because one comparison https://github.com/rvlenth/emmeans/issues/31
                export = TRUE) 

## Merge contrast tables for corn, soy and oat yields 
crop_ct <- rbind(corn_ct$summary, soy_ct$summary, oat_ct$summary)
crop_ct_df <- as.data.frame(crop_ct)

#crop_ct_df$contrast <- gsub("-", "vs", crop_ct_df$contrast )

crop_jt_ct <- cbind(crop_df, crop_ct_df[,-c(3:6)])

rownames(crop_jt_ct) <- NULL

crop_jt_ct_df <- as.data.frame(crop_jt_ct)
 
crop_jt_ct_df$`model term                  ` <- gsub("_", " ", crop_jt_ct_df$`model term                  `)
crop_jt_ct_df$`model term                  ` <- gsub(":", " x ", crop_jt_ct_df$`model term                  `)

colnames(crop_jt_ct_df) <- c("Source of variation", "df1", "df2","F","p", "contrast", "ratio", "p")


crop_jt_ct_df %>%
  kbl(booktabs = TRUE, linesep="",  align = "lrrrrlrr", digits = 2,
    caption = "Contrasts of rotation effect (expressed by Crop ID) on crop yields. The abbreviations on the contrast column are crop identities, which are the combinations of the first letter in crop species names and the rotation in which it occurred.") %>%
  column_spec(5, border_left = FALSE, border_right = TRUE) %>%
  add_header_above(c("ANOVA" = 5, "Comparison" = 3)) %>%
  pack_rows("(A) - Corn", 1, 3) %>%
  pack_rows("(B) - Soybean", 4, 6) %>%
  pack_rows("(C) - Oat", 7, 7) %>%
  footnote(general = "Corn weed management: low herbicide or conventional. Crop ID: crop species and the cropping system in which it occurred: C2 - corn in the 2-year rotation, C3 - corn in the 3-year rotation, C4 - corn in the 4-year rotation, S2 - soybean in the 2-year rotation, S3 - soybean in the 3-year rotation, S4 - soybean in the 4-year rotation, O3 - oat in the 3-year rotation, and O4 - oat in the 4-year rotation.", threeparttable = TRUE, footnote_as_chunk = TRUE)
```

```{r}
corn.emmip.bu <- emmip(corn.lmer.bu.ln, ~ Crop_ID,
                    CIs = TRUE,
                    type = "response", 
                    plotit = FALSE)

soy.emmip.bu <- emmip(soy.lmer.bu.ln, ~ Crop_ID,
                   CIs = TRUE, 
                   type = "response", 
                   plotit = FALSE)

oat.emmip.bu <- emmip(oat.lmer.bu.ln, ~ Crop_ID,
                   CIs = TRUE, 
                   type = "response",
                   plotit = FALSE)
```

```{r, include=FALSE}
## Prepare data sets for crop yield bar graphs as faceted single picture

#emmip gives the same numbers as emm, but emmip is tidier.
corn.emmip <- emmip(corn.lmer.ln, ~ Crop_ID,
                    CIs = TRUE,
                    type = "response", 
                    plotit = FALSE)

soy.emmip <- emmip(soy.lmer.ln, ~ Crop_ID,
                   CIs = TRUE, 
                   type = "response", 
                   plotit = FALSE)

oat.emmip <- emmip(oat.lmer.ln, ~ Crop_ID,
                   CIs = TRUE, 
                   type = "response",
                   plotit = FALSE)

alf.emmip <- emmip(alf.lmer.ln, ~1,
                   CIs = TRUE, 
                   type = "response",
                   plotit = FALSE)

names(alf.emmip) <- names(oat.emmip)
alf.emmip$Crop_ID <- gsub("overall", "A4", alf.emmip$Crop_ID)
alf.emmip$xvar <- gsub("overall", "A4", alf.emmip$xvar)

## All crop yields in one table 
crops.emmip <- rbind(corn.emmip, soy.emmip, oat.emmip, alf.emmip) 

## Duplicate the Crop_ID and save as character
crops.emmip$Crop_ID_char <- as.character(crops.emmip$Crop_ID)

## Add crop species name to merge with Iowa yields
crops.emmip_commodity <- crops.emmip %>%
  mutate(Commodity = ifelse(startsWith(Crop_ID_char,"S"),"soybeans",
                                        ifelse(startsWith(Crop_ID_char,"O"),"oats",
                                            ifelse(startsWith(Crop_ID_char,"C"),"corn","hay"))))%>%
    mutate(Rotation = ifelse(endsWith(Crop_ID_char,"4"),"4-year",
                                        ifelse(endsWith(Crop_ID_char,"3"),"3-year","2-year"))) %>%
  mutate(unit = ifelse(yvar > 50, "bu/ac", "ton/ac")) # id bu/ac
```


```{r, include=FALSE}
## Iowa and Boone yields 
IA_yield <- read.csv("../2-Data/Clean/NASS.yield.csv")

##Iowa average, need update when 2020 corn, 2019 and 2020 hay, 2019 oat, and 2020 soy yields are  available 
IA_avg_yield <- IA_yield %>% 
  group_by(Commodity) %>%
  summarise(IA_avg_ton = mean(Value.ton.ac, na.rm = TRUE),
            IA_sd_ton = sd(Value.ton.ac, na.rm = TRUE)) %>%
  mutate(IA_avg_bu = ifelse(Commodity == "corn", IA_avg_ton * 2000/ 56,
                                ifelse(Commodity == "soybeans", IA_avg_ton * 2000/ 60,
                                       ifelse(Commodity == "oats", IA_avg_ton * 2000/ 32 ,
                                              NA )))) 

Boone_yield <- IA_yield %>%
  filter(County == "BOONE") %>%
  group_by(Commodity) %>%
  summarise(Boone_avg_ton = mean(Value.ton.ac, na.rm = TRUE),
            Boone_sd_ton = sd(Value.ton.ac, na.rm = TRUE)) %>%
  mutate(Boone_avg_bu = ifelse(Commodity == "corn", Boone_avg_ton * 2000/ 56,
                                ifelse(Commodity == "soybeans", Boone_avg_ton * 2000/ 60,
                                       ifelse(Commodity == "oats", Boone_avg_ton * 2000/ 32,
                                              NA )))) 
```

```{r, include=FALSE}
## merge Marsden and Iowa/Boone yields 
crops.emmip.IA <- crops.emmip_commodity %>% 
  left_join(IA_avg_yield, by = "Commodity") %>%
  left_join(Boone_yield, by = "Commodity") %>%
  mutate(Commodity = recode(Commodity, "corn" = "corn grain",
                            "soybeans" = "soybean grain",
                            "oats" = "oat grain",
                            "hay" = "alfalfa hay")) 
# facet label name change: https://r-graphics.org/recipe-facet-label-text

## rearrange crop corresponding to their sequence in rotations
crops.emmip.IA$Commodity <- factor(crops.emmip.IA$Commodity, 
                                         levels=c("corn grain", 
                                                  "soybean grain",
                                                  "oat grain", 
                                                  "alfalfa hay"))
```

##### Yield table for extension communication

```{r}
corn.ext <- emmip(corn.lmer.bu.ln , ~ Crop_ID | Corn_weed_management,
                    CIs = TRUE,
                    type = "response", 
                    plotit = FALSE)

soy.ext  <- emmip(soy.lmer.bu.ln , ~ Crop_ID | Corn_weed_management,
                   CIs = TRUE, 
                   type = "response", 
                   plotit = FALSE)

oat.ext  <- emmip(oat.lmer.bu.ln , ~ Crop_ID,
                   CIs = TRUE, 
                   type = "response",
                   plotit = FALSE) %>% 
  mutate(Corn_weed_management = "both")

alf.ext <- alf.emmip %>% 
  mutate(Corn_weed_management = "both")
 
rbind(corn.ext, soy.ext, oat.ext , alf.ext) %>%
  select(Crop_ID, Corn_weed_management, yvar,
         LCL, UCL) %>%
  mutate(unit = ifelse(Crop_ID == "A4", "ton/ac", "bu/ac")) %>%
   mutate(across(where(is.numeric), ~ as.numeric(sprintf('%.1f', .)))) %>%
  mutate(`Yield [95% CI]` = paste0(yvar, " [", LCL, " , ", UCL, "]")) %>%
  select(Crop_ID, Corn_weed_management, `Yield [95% CI]`, unit) %>%
  arrange(Corn_weed_management)
```

```{r crop-bar-ton, fig.cap="Mean crop yields by rotation from 2017 to 2020. The color-coded bars show crop yields (ton ac$^-1$) in the experiment plots. The error bars show the 95% confidence intervals. The solid horizontal lines show mean yields for Iowa and dashed lines show mean yields for Boone County. Corn, soybean, and alfalfa yields in the experiment were averaged over four years, oat grain yields in the experiment were averaged over 2017, 2019, and 2020 because in 2018 oat was harvested for hay. Because county-specific alfalfa hay yields in 2019 and 2020 were unavailable at this writing, Boone County alfalfa yield (solid line) was averaged over 2017 and 2018 and Iowa hay yield was averaged over all counties in 2017 and 2018 and Iowa hay yield (dashed line) was averaged from all county-based values in 2017 and 2018 and two state-based values in 2019 and 2020. Corn, soybean, oat grain, and alfalfa hay yields were reported at 15.5%, 13%, 14%, and 0% moisture respectively.", echo=FALSE, fig.width=8, fig.height=6}

crops.emmip.IA %>%
  filter(unit == "ton/ac") %>%
ggplot(aes(x = Crop_ID, y = yvar, fill = Rotation))+ 
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Dark2") +
  geom_errorbar(aes(ymin= LCL, ymax= UCL), width=.2) + 
  facet_grid(.~ Commodity, scales = "free", space = "free") + # scales: remove empty bar, space: equal bar width
  geom_hline(aes(yintercept = IA_avg_ton, linetype = "IA average")) + 
  geom_hline(aes(yintercept = Boone_avg_ton, linetype = "Boone County average")) + 
  ylab("Crop yield (ton/ac)") +
  xlab("Crop identity") +
  labs(linetype = "Yield") +
  theme(legend.position="bottom") 
```


