---
output: 
  # bookdown::pdf_document2: 
  #   keep_tex: yes
  #   toc: false
  #   number_sections: false
  #  extra_dependencies: ["float"]
  bookdown::html_document2:
    toc: false
    fig_caption: yes
    keep_md: true
    self_contained: true
---


```{r,  echo=FALSE, message=FALSE, warning=FALSE}
#install.packages("here")
library(emmeans)
library(here)
library(readr)
library(data.table)
#need relative path for figures
knitr::opts_chunk$set(message = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                     # fig.width = 8, #change number specs to balance plot size and resolution
                     # fig.height = 8,
                     # fig.fullwidth = TRUE,
                     # fig.path = here("4 Graphs/"), 
                      dev=c("png"),
                      fig.pos = "H", 
                      table.placement = "H")
                   #   out.extra = "")

library(tidyverse)
library(magrittr)
library(lme4)
library(kableExtra)
theme_set(theme_bw())
library(ggResidpanel)
library(gridExtra)
```


```{r, include=FALSE}
dens_1720 <- read.csv(here("2-Data/Clean/pldens_1720_clean.csv"))
dens_1720_clean <- dens_1720 %>% replace(is.na(.), 0) %>%
  mutate(zOTHERS = Total - (AMATA + CHEAL + DIGSA + ECHCG + SETFA + SETLU + TAROF)) %>%#keep top seven species and sum all other species in to zOTHERS
  dplyr::select(Crop, Crop_ID, Block, Year, Corn_weed_management,
    AMATA , CHEAL , DIGSA , ECHCG , SETFA , SETLU , TAROF, zOTHERS) %>%
  mutate_at(13, funs(round(., 8))) # zOTHERS make the group listed last in ggplot2



biom_1720 <- read.csv(here("2-Data/Clean/biom_1720_clean.csv"))
biom_1720_clean <- biom_1720 %>% replace(is.na(.), 0)%>%
   mutate(zOTHERS = Total - (AMATA + CHEAL + DIGSA + ECHCG + SETFA + SETLU + TAROF)) %>%#keep top seven species and sum all other species in to zOTHERS
  dplyr::select(Crop, Crop_ID, Block, Year, Corn_weed_management,
                AMATA , CHEAL , DIGSA , ECHCG , SETFA , SETLU , TAROF, zOTHERS) %>%
  mutate_at(13, funs(round(., 8))) # zOTHERS make the group listed last in ggplot2


# remove rows that sum to zeros https://www.tutorialspoint.com/how-to-remove-rows-that-contains-all-zeros-in-an-r-data-frame
#dens_1720_clean <- dens_1720[rowSums(dens_1720[,13:46])>0,] 

dens_1720_clean$Crop <- factor(dens_1720_clean$Crop,
                               levels = c("corn", "soybean", "oat", "alfalfa"))
dens_1720_clean$Crop_ID <- factor(dens_1720_clean$Crop_ID,
                                  levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

#biom_1720_clean <- biom_1720[rowSums(biom_1720[,13:46])>0,] 
biom_1720_clean$Crop <- factor(biom_1720_clean$Crop, 
                               levels = c("corn", "soybean", "oat", "alfalfa"))

biom_1720_clean$Crop_ID <- factor(biom_1720_clean$Crop_ID,
                                  levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

dens_1720_clean %<>% 
  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

biom_1720_clean %<>% 
  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

# Find minimum non-zero in each matrix, or in the Prop column of the long-format dframe  
```


```{r}
dens_1720_clean %>%
  group_by(Corn_weed_management, Crop_ID, Year) %>%
  summarise(meanP_pl = mean(AMATA)) %>%
  filter(Year %in% c("2018", "2019"), Corn_weed_management == "conv")
```

```{r, include=FALSE}

##### Fit all species at once with split-fit-combine syntax 
# https://stat585-at-isu.github.io/materials-2019/04_functional-programming/02_purrr.html#23

## contrast coefficients for hand-written contrasts 

C2 <- c(1,rep(0,8)); C3 <- c(0,0,1,rep(0,6)); C4 <- c(rep(0,5),1,rep(0,3))
S2 <- c(0, 1,rep(0,7)); S3 <- c(rep(0,3), 1,rep(0,5)); S4 <- c(rep(0,6),1,rep(0,2))
O3 <- c(rep(0,4), 1, rep(0,4)); O4 <- c(rep(0,7),1,0)
A4 <- c(rep(0,8), 1)


# convert wide to long format 

dens_1720_long <- dens_1720_clean %>% 
  pivot_longer(!c(Crop:Corn_weed_management),
               names_to = "Species",
               values_to =  "Density")

# justification of betareg https://www.sciencedirect.com/science/article/pii/S1470160X16305659#sec0020
# multinomial logistic seems more reasonable https://besjournals.onlinelibrary.wiley.com/doi/10.1111/2041-210X.13234
# https://github.com/bobdouma/proportions_beta_Dirichlet/blob/v.01/Appendix%20S4%20Case_study%202.Rmd
#https://orcid.org/0000-0002-8779-838X
#https://stats.stackexchange.com/questions/504644/beta-regression-fitted-values

#singularity because of too many zeros in many species columns 


min(dens_1720_long$Density[dens_1720_long$Density>0]) #0.05396072

dens_result_t <- dens_1720_long %>%
  group_by(Species) %>% # to fit a linear model for each species
  nest() %>%
  mutate(models=map(data,~lmer(log(Density + 0.05396072) ~ Block 
                               + Crop_ID + 
                     Corn_weed_management 
                     + Crop_ID:Corn_weed_management + 
                        (1|Year) 
                     + (1|Year:Block) + 
                       (1|Year:Crop_ID) 
                     + (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  +
                       (1|Block:Year:Crop_ID) ,
                     data =.x))) %>%
  ungroup %>%
  mutate(diag_plots = map(models, resid_panel),
    jts = map(models, joint_tests), #ANOVA table, with Block effect
    jt_cleans = map2(data, jts, ~print(.y, export = TRUE)$summary), #ANOVA tables with Block effect in clean format
         means = map2(data, models,~emmeans(.y, ~ Crop_ID|Corn_weed_management,data=.x, type = "response")), #estimated mean density of each weed species in each crop ID, separated by corn weed management
        means2 = map2(data, models,~emmeans(.y,  c("Crop_ID","Corn_weed_management"),data=.x, type = "response")), # EM means, for ANOVA table, averaged over block
    means_crop = map2(data, models,~emmeans(.y,  "Crop_ID",data=.x, type = "response")), #EM means averaged by Crop id
    means_herb = map2(data, models,~emmeans(.y,  "Corn_weed_management",data=.x, type = "response")), #EM means averaged by corn weed mgt
     jts2 = map(means2, joint_tests), #ANOVA table, averaged over block
    jts2_clean = map2(data, jts2, ~print(.y, export = TRUE)$summary), #ANOVA table, averaged over block in clean format
     arrow_plots = map(means, plot, comparison = TRUE, xlab=expression(plantsm^{"-2"}), ylab = "Crop ID"), #visualize EM means of stand density

dens_2yrvs34yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+O3+C4+S4+O4+A4)/7]" = ((C2+S2)/2) - ((C3+S3+O3+C4+S4+O4+A4)/7) )),
dens_3yrvs4yr = map(means_crop, contrast, method = list("[(C3+S3+O3)/3] vs [(C4+S4+O4+A4)/4]" = ((C3+S3+O3)/3) - ((C4+S4+O4+A4)/4) )),
dens_2yrvs3yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+O3)/3]" = ((C2+S2)/2) -  ((C3+S3+O3)/3) )),
dens_2yrvs4yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C4+S4+O4+A4)/4]" = ((C2+S2)/2) - ((C4+S4+O4+A4)/4) )),
dens_CS2yr_vs_CS34yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+C4+S4)/4]" = ((C2+S2)/2) - ((C3+S3+C4+S4)/4))),
dens_CS3yr_vs_CS4yr = map(means_crop, contrast, method = list("[(C3+S3)/2] vs [(C4+S4)/2]" = ((C3+S3)/2) - ((C4+S4)/2))),
dens_CS2yr_vs_CS3yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3)/2]" = ((C2+S2)/2) - ((C3+S3)/2))),
dens_CS2yr_vs_CS4yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C4+S4)/2]" = ((C2+S2)/2) - ((C4+S4)/2))),
dens_C2vs3_4 = map(means_crop, contrast, method = list("C2 vs [(C3+C4)/2]" = C2 - ((C3+C4)/2) )),
dens_C2vs4 = map(means_crop, contrast, method = list("C2 vs C4" = C2 - C4)),
dens_C2vs3 = map(means_crop, contrast, method = list("C2 vs C3" = C2 - C3)),
dens_C3vs4 = map(means_crop, contrast, method = list("C3 vs C4" = C3 - C4)),
dens_S2vs3_4 = map(means_crop, contrast ,method = list("S2 vs [(S3+S4)/2]" = S2 - ((S3+S4)/2) )),
dens_S2vs4 = map(means_crop, contrast, method = list("S2 vs S4" = S2 - S4)),
dens_S2vs3 = map(means_crop, contrast, method = list("S2 vs S3" = S2 - S3)),
dens_S3vs4 = map(means_crop, contrast, method = list("S3 vs S4" = S3 - S4)),
dens_oat = map(means_crop,contrast, method = list("O3 vs O4" = O3 - O4 )),
dens_summer_vs_cool = map(means_crop, contrast,  method = list("[(O3+O4+A4)/3] vs [(C2+S2+C3+S3+C4+S4)/6]" = (((O3+A4+O4)/3)) - (C2+S2+C3+S3+C4+S4)/6)),
dens_CS3_vs_O3 = map(means_crop, contrast, method = list("O3 vs [(C3+S3)/2]" = O3 - ((C3+S3)/2))),
dens_CS4_vs_OA4 = map(means_crop, contrast, method = list("[(O4+A4)/2] vs [(C4+S4)/2]" = ((A4+O4)/2) - ((C4+S4)/2))),
dens_forage3_vs4 = map(means_crop, contrast,  method = list("[(O3+O4)/2] vs A4" = ( ((O3+O4)/2)) - A4)), 
dens_conv_low = map(means_herb, contrast,  "pairwise"), 
         p_cont = map(means, contrast, "pairwise"), 
         across(models:p_cont, setNames,  .$Species))


biom_1720_long <- biom_1720_clean %>% 
  pivot_longer(!c(Crop:Corn_weed_management), names_to = "Species", values_to =  "Biomass")

min(biom_1720_long$Biomass[biom_1720_long$Biomass>0]) # 0.0005396072

biom_result_t <- biom_1720_long %>%
  group_by(Species) %>% # to fit a linear model for each species
  nest() %>%
  mutate(models=map(data,~lmer(log(Biomass + 0.0005396072) ~ Block +
                                 Crop_ID*Corn_weed_management + 
                        (1|Year) + (1|Year:Block) + 
                       (1|Year:Crop_ID) + (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  + (1|Block:Year:Crop_ID),data =.x))) %>%
  ungroup %>%
  mutate(jts = map(models, joint_tests), #ANOVA table, with Block effect
         diag_plots = map(models, resid_panel),  #diagnosis plot
    jt_cleans = map2(data, jts, ~print(.y, export = TRUE)$summary), #ANOVA table, with Block effect with clean format
         means = map2(data, models,~ emmeans(.y, ~ Crop_ID | Corn_weed_management,data=.x, type = "response")), #estimated mean density of each weed species in each crop ID, separated by corn weed management, for arrow_plot
        means2 = map2(data, models,~emmeans(.y,  c("Crop_ID","Corn_weed_management"),data=.x, type = "response")), # EM means, for ANOVA table, averaged over block
    jts2 = map(means2, joint_tests), #ANOVA table, averaged over block 
    jts2_clean = map2(data, jts2, ~print(.y, export = TRUE)$summary), #ANOVA table, averaged over block in clean format
        means_crop = map2(data, models,~emmeans(.y,  "Crop_ID",data=.x, type = "response")), # EM means of weed aboveground mass in Crop ID, averaged over herbicide
    means_herb = map2(data, models,~emmeans(.y,  "Corn_weed_management",data=.x, type = "response")),  # EM means of weed aboveground mass in herbicide, averaged over crop ID
    arrow_plots = map(means, plot, comparison = TRUE, xlab=expression(gm^{"-2"}), ylab = "Crop ID"),  #visualize EM means of aboveground mass
biom_2yrvs34yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+O3+C4+S4+O4+A4)/7]" = ((C2+S2)/2) - ((C3+S3+O3+C4+S4+O4+A4)/7) )),
biom_3yrvs4yr = map(means_crop, contrast, method = list("[(C3+S3+O3)/3] vs [(C4+S4+O4+A4)/4]" = ((C3+S3+O3)/3) - ((C4+S4+O4+A4)/4) )),
biom_2yrvs3yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+O3)/3]" = ((C2+S2)/2) -  ((C3+S3+O3)/3) )),
biom_2yrvs4yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C4+S4+O4+A4)/4]" = ((C2+S2)/2) - ((C4+S4+O4+A4)/4) )),
biom_CS2yr_vs_CS34yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3+C4+S4)/4]" = ((C2+S2)/2) - ((C3+S3+C4+S4)/4))),
biom_CS3yr_vs_CS4yr = map(means_crop, contrast, method = list("[(C3+S3)/2] vs [(C4+S4)/2]" = ((C3+S3)/2) - ((C4+S4)/2))),
biom_CS2yr_vs_CS3yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C3+S3)/2]" = ((C2+S2)/2) - ((C3+S3)/2))),
biom_CS2yr_vs_CS4yr = map(means_crop, contrast, method = list("[(C2+S2)/2] vs [(C4+S4)/2]" = ((C2+S2)/2) - ((C4+S4)/2))),
biom_C2vs3_4 = map(means_crop, contrast, method = list("C2 vs [(C3+C4)/2]" = C2 - ((C3+C4)/2) )),
biom_C2vs4 = map(means_crop, contrast, method = list("C2 vs C4" = C2 - C4)),
biom_C2vs3 = map(means_crop, contrast, method = list("C2 vs C3" = C2 - C3)),
biom_C3vs4 = map(means_crop, contrast, method = list("C3 vs C4" = C3 - C4)),
biom_S2vs3_4 = map(means_crop, contrast ,method = list("S2 vs [(S3+S4)/2]" = S2 - ((S3+S4)/2) )),
biom_S2vs4 = map(means_crop, contrast, method = list("S2 vs S4" = S2 - S4)),
biom_S2vs3 = map(means_crop, contrast, method = list("S2 vs S3" = S2 - S3)),
biom_S3vs4 = map(means_crop, contrast, method = list("S3 vs S4" = S3 - S4)),
biom_oat = map(means_crop,contrast, method = list("O3 vs O4" = O3 - O4 )),
biom_summer_vs_cool = map(means_crop, contrast,  method = list("[(O3+O4+A4)/3] vs [(C2+S2+C3+S3+C4+S4)/6]" = (((O3+A4+O4)/3)) - (C2+S2+C3+S3+C4+S4)/6)),
biom_CS3_vs_O3 = map(means_crop, contrast, method = list("O3 vs [(C3+S3)/2]" = O3 - ((C3+S3)/2))),
biom_CS4_vs_OA4 = map(means_crop, contrast, method = list("[(O4+A4)/2] vs [(C4+S4)/2]" = ((A4+O4)/2) - ((C4+S4)/2))),
biom_forage3_vs4 = map(means_crop, contrast,  method = list("[(O3+O4)/2] vs A4" = ( ((O3+O4)/2)) - A4)),
biom_conv_low = map(means_herb, contrast,  "pairwise"),
p_cont = map(means, contrast, "pairwise"), 
across(models:p_cont, setNames,  .$Species))

```

#### How did rotation, crop species, and corn weed management affect individual weed species abundance?{-}  

*The hypothesis that "including oat and alfalfa in rotations with corn and soybean will reduce the density and aboveground mass of noxious weed species in corn and soybean" was partially supported.* Crop identity affected individual density of seven most abundant weed species but corn weed management affected that of two weed species only, i.e., DIGSA and SETFA (p = 0.0189 and p = 0.0196, resepectively; Table \@ref(tab:ind-dens-biom-jt). Among those seven weed species, the aboveground mass of four (CHEAL, DIGSA, SETFA, and TAROF) were affected by crop identity, but none was affected by corn weed management (Table \@ref(tab:ind-dens-biom-jt). The magnitude of difference in stand density and aboveground mass were the most pronounced between crop types (Table \@ref(tab:indiv-dens-biom-ct)). The main-plot effects concerning crop identity on individual species responses are elaborated below.  


```{r ind-dens-biom-jt, echo=FALSE}
#separate by density and biomass because of large table 

ind_dens_jt <- rbind(dens_result_t$jts2_clean$AMATA,
                     dens_result_t$jts2_clean$CHEAL,
                     dens_result_t$jts2_clean$DIGSA,
                     dens_result_t$jts2_clean$ECHCG,
                     dens_result_t$jts2_clean$SETFA,
                     dens_result_t$jts2_clean$SETLU,
                     dens_result_t$jts2_clean$TAROF, 
  dens_result_t$jts2_clean$zOTHERS) 
ind_dens_df <- as.data.frame(ind_dens_jt)


ind_dens_df$F.ratio <- as.numeric(ind_dens_df$F.ratio)

ind_biom_jt <- rbind(biom_result_t$jts2_clean$AMATA,
      biom_result_t$jts2_clean$CHEAL,
      biom_result_t$jts2_clean$DIGSA,
      biom_result_t$jts2_clean$ECHCG,
      biom_result_t$jts2_clean$SETFA,
      biom_result_t$jts2_clean$SETLU,
      biom_result_t$jts2_clean$TAROF,
      biom_result_t$jts2_clean$zOTHERS) 

ind_biom_df <- as.data.frame(ind_biom_jt)
ind_biom_df$F.ratio <- as.numeric(ind_biom_df$F.ratio)


ind_dens_biom_jt <- cbind(ind_dens_df, ind_biom_df)
# can't do left_join here because rows will be reorganized alphabetically

# remove "_" and replace ":" with "x"
ind_dens_biom_jt$`model term                  ` <- gsub("_", " ", ind_dens_biom_jt$`model term                  `)
ind_dens_biom_jt$`model term                  ` <- gsub(":", " x ", ind_dens_biom_jt$`model term                  `)

rownames(ind_dens_biom_jt) <- NULL

ind_dens_biom_jt_clean <- ind_dens_biom_jt[,-c(6,7,8)] 
colnames(ind_dens_biom_jt_clean) <- c("Source of variation", "df1", "df2", rep(c("F","p"),2))

ind_dens_biom_jt_clean %>%
  kbl(booktabs = TRUE, 
      digits = 2,
      linesep = " ", longtable = FALSE, align = "lrrrrrr",
    caption = "Treatment effects on the stand density and aboveground mass of the seven most abundant weed species, listed alphabetically. All the other weeds species were grouped into OTHERS.") %>% 
  add_header_above(c(" " = 3, "Stand density" = 2, "Aboveground mass" = 2)) %>% 
  pack_rows("(A) - AMATA", 1, 3) %>%
  pack_rows("(B) - CHEAL", 4, 6) %>% 
    pack_rows("(C) - DIGSA", 7, 9) %>% 
  pack_rows("(D) - ECHCG",  10, 12) %>%
    pack_rows("(E) - SETFA",13, 15) %>% 
  pack_rows("(F) - SETLU", 16, 18) %>% 
  pack_rows("(G) - TAROF", 19, 21) %>% 
  pack_rows("(H) - OTHERS", 22, 24) %>% 
    column_spec(5,border_left = FALSE,
                border_right = TRUE) %>%
  kable_styling(latex_options = c("scale_down")) %>%
  footnote(general = "Corn weed management: low herbicide or conventional. C2 - corn in the 2-year rotation, C3 - corn in the 3-year rotation, C4 - corn in the 4-year rotation, S2 - soybean in the 2-year rotation, S3 - soybean in the 3-year rotation, S4 - soybean in the 4-year rotation, O3 - oat in the 3-year rotation, O4 - oat in the 4-year rotation, and A4 - alfalfa in the 4-year rotation.", threeparttable = TRUE, footnote_as_chunk = TRUE)
```

```{r}
biom_result_t$means %>%
  map(~ broom::tidy(.x)) %>%
 bind_rows(.id = "spp") %>%
  select(1:5) 
```

*The cool-season crops were responsible for AMATA stand density differences, but those differences were not strong enough to be apparent between rotation averages.* AMATA stand density and aboveground mass were comparable among all rotation systems averaged over crop phases (p-values > 0.05), among rotations for the same crop species (p-values > 0.05), and within the same crop type across rotations (p-values > 0.05). Averaged over the same crop types (warm-season or cool-season), AMATA stand density in cool-season was 12.25-fold greater than that in warm-season crops (p = 0.0001), but AMATA aboveground mass was comparable in cool-season and warm-season crops (p = 0.0906). Within the same rotation, AMATA stand density was 11-fold (p = 0.0143) and 23-fold (p = 0.0003) greater in the cool-season than in the warm-season crops overall averages, but AMATA aboveground mass was comparable in these crop environments (p = 0.2355 and p = 0.0493, respectively).  


*The cool-season crops, especially oat were responsible for CHEAL stand density and aboveground mass differences between rotation averages.* CHEAL stand density and aboveground mass were 4-fold (p = 0.008) and 5-fold (p = 0.199) greater in the average of the 3-year and 4-year rotations than in the 2-year rotation, but comparable between the 3-year and 4-year rotations (p = 0.9195 and p = 0.6114, respectively). CHEAL stand density and aboveground mass were comparable between rotations for the same crop species (p-values > 0.05) and within the warm-season crops (p-values > 0.05). CHEAL stand density and aboveground mass were 38-fold (p < 0.0001) and 204-fold (p < 0.0001) greater in the cool-season crops than in the warm-season crops overall averages; 67-fold (p < 0.0001) and 571-fold (p < 0.0001) greater in the cool-season crop than in the warm-season crops average of the 3-year rotation; and 37-fold (p < 0.0001) and 232-fold (p < 0.0001) greater in the cool-season crop than in the warm-season crops average of the 4-year rotation. CHEAL stand density and aboveground mass were 11-fold (p = 0.0001) and 96-fold (p = 0.0001) greater in oat than in alfalfa. 


*The cool-season crops, especially alfalfa were responsible for DIGSA stand density and aboveground mass differences between rotation averages.* DIGSA stand density in the average of the 3-year and 4-year rotations was two-fold greater than in the 2-year rotation (p = 0.0072) and 5-fold greater in the 4-year rotation than in the 3-year rotation (p < 0.0001). DIGSA aboveground mass was comparable between the 2-year and the average of the 3-year and 4-year rotations (p = 0.1098), but 14-fold greater in the 4-year than in the 3-year rotations (p = 0.0001). DIGSA stand density and aboveground mass were comparable between rotations for the same crop species (p-values > 0.05), except for oat (p = 0.0062 and p = 0.0032). DIGSA stand density and aboveground mass were 10-fold and 27-fold greater in the cool-season crop averages than in the warm-season crops averages, 20-fold (p = 0.0001) and 103-fold (p = 0.0001) greater in the cool-season crops than in the warm-season crops of the 4-year rotation, but comparable between cool-season and warm-season crops of the 3-year rotation (p = 0.0603 and p = 0.3924, respectively). DIGSA stand density and aboveground mass were 14-fold (p = 0.0001) and 33-fold (p = 0.0001) greater in alfalfa than in oat.   


*ECHCG responses generally were similar to those of AMATA.* ECHCG stand density and aboveground mass were comparable between all rotation averages (p-values > 0.05), between rotations for the same crop species (p-values > 0.05), within the same crop type between rotations (p-values > 0.05), and within the 3-year rotation (p-values > 0.05). Averaged over the same crop types, ECHCG stand density and aboveground mass were 4-fold (p = 0.0003) and 10-fold (p = 0.0012) greater in the cool-season than in the warm-season crops. Within the 4-year rotation, ECHCG stand density and aboveground mass were 5-fold (p = 0.0014) and 18-fold (p = 0.0031) greater in the cool-season than in the warm-season crops.   


*The cool-season crops were responsible for SETFA stand density and aboveground mass differences, but those differences were not strong enough be apparent between rotation averages.* SETFA stand density and aboveground mass were comparable between all rotation averages (p-values > 0.05), between rotations for the same crop species (p-values > 0.05), within the warm-season crops between rotations (p-values > 0.05), and within the cool-season crops (p-values > 0.05). Averaged over the same crop types, SETFA stand density and aboveground mass were 10-fold (p < 0.0001) and 15-fold (p = 0.0008) greater in the cool-season than in the warm-season crops. Within the same rotation, SETFA stand density and aboveground mass were 11-fold to 23-fold greater in the cool-season than in the warm-season crops (Table \@ref(tab:indiv-dens-biom-ct)).  

SETLU stand density and aboveground mass were comparable in most pairs of comparison (p-values > 0.05), except that SETLU stand density was 2.5-fold greater in the cool-season crops average than in the warm-season crops average(p = 0.0404).  

*The cool-season crops, especially oat were responsible for TAROF stand density and aboveground mass differences between rotation averages.* TAROF stand density and aboveground mass in the 3-year and 4-year rotations average were 4-fold (p < 0.0001) and 14-fold (p < 0.0001) greater than those in the 2-year rotation. TAROF stand density and aboveground mass in the 3-year rotation were and 5-fold  (p < 0.0001) and 20-fold (p < 0.0001) greater than those in the 4-year rotation. TAROF stand density and aboveground mass were comparable among the warm-season crops between rotations and within the same crops between rotations (p-values > 0.05), except in oat (p < 0.0001). TAROF stand density and aboveground mass were 24-fold (p < 0.001) and 390-fold (p < 0.0001) greater in cool-season than in warm-season crop averages, 4-fold (p = 0.0001) and 20-fold (p = 0.0002) greater in oat than in corn and soybean averages in the 3-year rotation, and 54-fold (p < 0.0001) and 1483-fold (p < 0.0001) greater in the cool-season crops than in the warm-season crops in the 4-year rotation. TAROF stand density and aboveground mass were 6-fold (p < 0.0001) and 20-fold (p = 0.0001) greater in oat than in alfalfa.


```{r, include=FALSE}

## select string starts with "dens_
dens_ct <- dens_result_t %>% select(contains("dens_")) #extract the contrasts only

## need something here to preserve p-value formats
dens_ct_list_df <- map(dens_ct , as.data.frame) # same as lapply(dens_ct, as.data.frame, stringAsFactors = FALSE) #give a list of 32 df of 1x238, each df is a contrast


dens_ct_df <- lapply(dens_ct_list_df, 
                     as.data.frame, stringsAsFactors = FALSE) %>%
  rbindlist(fill = FALSE) #combine everything into a large df
```

```{r, include=FALSE}
# Keep contrast ratio and p-values in density 
dens_ct_df_clean <- dens_ct_df %>%
  select(-contains(c(".SE", ".df", ".null", ".t.ratio", "OTHER")))  %>%
  select(-contains(c("CHEAL.contrast", "DIGSA.contrast", "ECHCG.contrast", "SETFA.contrast","SETLU.contrast", "TAROF.contrast")))


dens_ct_df_rounded <- dens_ct_df_clean %>%
  mutate_at(c(2,4,6,8,10,12,14), ~ sprintf('%.2f', .)) %>% #all ratios 2 digits
  mutate_at(c(3,5,7,9,11,13,15), ~ sprintf('%.4f', .)) %>% #all p-values at 4 digits
  mutate(across(0.0000, ~ as.character)) %>% 
  mutate_if(is.character, str_replace_all,
            pattern = '0.0000', 
            replacement = '<.0001')
```


```{r, include=FALSE}
## select string starts with "biom_
biom_ct <- biom_result_t %>% select(contains("biom_")) #extract the contrasts only

biom_ct_list_df <- map(biom_ct , as.data.frame) # same as lapply(dens_ct, as.data.frame, stringAsFactors = FALSE) #give a list of 32 df of 1x238, each df is a contrast


biom_ct_df <- lapply(biom_ct_list_df, 
                     as.data.frame,
                     stringsAsFactors = FALSE) %>%
  rbindlist(fill = FALSE) #combine everything into a large df
```



```{r, include=FALSE}
# Keep contrast ratio and p-values in density 
biom_ct_df_clean <- biom_ct_df %>%
  select(-contains(c(".SE", ".df", ".null", ".t.ratio", "OTHER")))  %>%
  select(-contains(c("CHEAL.contrast", "DIGSA.contrast", "ECHCG.contrast", "SETFA.contrast","SETLU.contrast", "TAROF.contrast")))


biom_ct_df_rounded <- biom_ct_df_clean %>%
  mutate_at(c(2,4,6,8,10,12,14), ~ sprintf('%.2f', .)) %>% #all ratios 2 digits
  mutate_at(c(3,5,7,9,11,13,15), ~ sprintf('%.4f', .)) %>% #all p-values at 4 digits
  mutate(across(0.0000, ~ as.character)) %>% 
  mutate_if(is.character, str_replace_all,
            pattern = '0.0000', 
            replacement = '<.0001')
```

##### Relative abundance of biomass 
```{r}
biom_1720_clean.raw <- biom_1720 %>% replace(is.na(.), 0)%>%
  mutate(SETSP = SETFA + SETLU) %>%
  dplyr::select( Crop_ID, Block, Year, Corn_weed_management, Total,
                AMATA , CHEAL , DIGSA , ECHCG , SETSP , TAROF) %>%
   mutate(
    AMATA_percent = (AMATA / Total) * 100,
    CHEAL_percent = (CHEAL / Total) * 100,
    DIGSA_percent = (DIGSA / Total) * 100,
    ECHCG_percent = (ECHCG / Total) * 100,
    SETSP_percent = (SETSP / Total) * 100,
    TAROF_percent = (TAROF / Total) * 100) %>%
    dplyr::select(Year, Block, 
                  Crop_ID, Corn_weed_management, 
                  AMATA_percent, CHEAL_percent, DIGSA_percent,
                  ECHCG_percent, SETSP_percent, TAROF_percent) 
```

```{r, include=FALSE, eval=FALSE}
biom_1720_clean.ext.raw <-  biom_1720_clean.raw %>%
   group_by(Crop_ID, Corn_weed_management) %>%  
  summarise(
    across(ends_with("_percent"), list(mean = ~ mean(.x, na.rm = TRUE), 
                                       sd = ~ sd(.x, na.rm = TRUE)), .names = "{.col}_{.fn}")
  ) %>%
  ungroup() %>%
  mutate(
    AMATA = paste0(round(AMATA_percent_mean, 1), " (", round(AMATA_percent_sd, 1), ")"),
    CHEAL = paste0(round(CHEAL_percent_mean, 1), " (", round(CHEAL_percent_sd, 1), ")"),
    DIGSA = paste0(round(DIGSA_percent_mean, 1), " (", round(DIGSA_percent_sd, 1), ")"),
    ECHCG = paste0(round(ECHCG_percent_mean, 1), " (", round(ECHCG_percent_sd, 1), ")"),
    SETSP = paste0(round(SETSP_percent_mean, 1), " (", round(SETSP_percent_sd, 1), ")"),
    TAROF = paste0(round(TAROF_percent_mean, 1), " (", round(TAROF_percent_sd, 1), ")")
  ) #%>%
  # Select the desired columns
  arrange(Corn_weed_management, Crop_ID)

```

```{r}
biom_1720_clean.raw_long <-  biom_1720_clean.raw %>%
  pivot_longer(!c(Year : Corn_weed_management),
               names_to = "Species",
               values_to =  "Relative.abundance")
```

```{r}
tran <- make.tran("asin.sqrt", 100)

indv_spp_result_t <- biom_1720_clean.raw_long %>%
  group_by(Species) %>% # to fit a linear model for each species
  nest() %>%
  mutate(models=map(data,~lmer(Relative.abundance ~ Block +
                                 Crop_ID*Corn_weed_management + 
                        (1|Year) + (1|Year:Block) + 
                       (1|Year:Crop_ID) + (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  + (1|Block:Year:Crop_ID),data =.x))) %>%
  mutate(models.t = map(data, ~with(tran, 
                                  lmer(linkfun(Relative.abundance) ~ Block +
                                         Crop_ID * Corn_weed_management + 
                                         (1|Year) + (1|Year:Block) + 
                                         (1|Year:Crop_ID) + (1|Year:Corn_weed_management) + 
                                         (1|Year:Crop_ID:Corn_weed_management) + (1|Block:Year:Crop_ID),
                                       data = .x)))) %>%
  ungroup %>%
  mutate(jts = map(models.t, joint_tests), #ANOVA table, with Block effect
         diag_plots = map(models, resid_panel),  #diagnosis plot
        diag_plots.t = map(models.t, resid_panel), 
    jt_cleans = map2(data, jts, ~print(.y, export = TRUE)$summary), #ANOVA table, with Block effect with clean format
         means = map2(data, models.t,~ emmeans(.y, ~ Crop_ID | Corn_weed_management,data=.x, type = "response")), #estimated mean density of each weed species in each crop ID, separated by corn weed management, for arrow_plot
        means2 = map2(data, models.t,~emmeans(.y,  c("Crop_ID","Corn_weed_management"),data=.x, type = "response")), # EM means, for ANOVA table, averaged over block
    jts2 = map(means2, joint_tests), #ANOVA table, averaged over block 
    jts2_clean = map2(data, jts2, ~print(.y, export = TRUE)$summary), #ANOVA table, averaged over block in clean format
across(models:jts2_clean, setNames,  .$Species))
```
Arsine squareroot improved model fitting.  

```{r}
indv_spp_result_t$diag_plots
```

```{r}
indv_spp_result_t$diag_plots.t
```

```{r}
indv_spp_result_t$means %>%
   map(~ broom::tidy(.x)) %>%
 bind_rows(.id = "spp") %>%
  select(1:5) %>%
    mutate(across(where(is.numeric), ~ as.numeric(sprintf('%.2f', .))*100)) %>%
  mutate(Relative.abundance = paste0(response, " (", std.error, ")")) %>%
  select(-c(response, std.error)) %>%
  mutate(spp = substr(spp, 1, 5)) %>%
  pivot_wider(names_from = spp, values_from = Relative.abundance) %>%
    mutate(Crop_ID = factor(Crop_ID, levels = c("C2", "C3", "C4",
                                                "S2", "S3", "S4",
                                                "O3", "O4", "A4"))) %>% # Replace with the specific order of Crop_ID you need
  arrange(Corn_weed_management, Crop_ID) #unknown error why the knitted table is totally irrelevant of this table
```

```{r, include=FALSE}
## stack density and aboveground mass, keep the rows with the most representative contrasts only 

dens_ct_condensed <- dens_ct_df_rounded[c(1,2,
                     5,6,
                     9,12,
                     13, 16, 17,
                     18:21),]

biom_ct_condensed <- biom_ct_df_rounded[c(1,2,
                     5,6,
                     9,12,
                     13, 16, 17,
                     18:21),]

dens_biom_ct_condensed <- rbind(dens_ct_condensed, biom_ct_condensed )

```

```{r indiv-dens-biom-ct, echo=FALSE}
dens_biom_ct_condensed  %>%
  `colnames<-` (c("Contrast of the main-plot effect", rep(c("ratio", "p"), 7))) %>%
    kbl(longtable = FALSE, booktabs = TRUE, linesep = "", align = 'lrrrrrrrrrrrrrr',
    caption = "Contrast of stand density and aboveground mass of the seven most abundant weed species. Weed species are listed alphabetically. The abbreviations on the contrast column are crop identities, which are the combinations of the first letter in crop species names and the rotation in which it occurred.") %>% 
  add_header_above(c(" " = 1, "AMATA" = 2, "CHEAL" = 2, "DIGSA" = 2, "ECHCG" = 2, "SETFA" = 2, "SETLU" = 2, "TAROF" = 2)) %>% 
    column_spec(c(3,5,7,9,11,13), border_left = FALSE, border_right = TRUE) %>%
   pack_rows("(A) - Stand density", 1, 13) %>%
  pack_rows("(A1) - Rotation system effects", 1, 4) %>%
  pack_rows("(A2) - Rotation system effects within individual crops", 5, 9) %>%
  pack_rows("(A3) - Crop type effects", 10, 13) %>%
  pack_rows("(B) - Aboveground mass", 14, 26) %>%
  pack_rows("(B1) - Rotation system effects", 14, 17) %>%
  pack_rows("(B2) - Rotation system effects within individual crops", 18, 22) %>%
  pack_rows("(B3) - Crop type effects", 23, 26) %>%
  kable_styling(latex_options = "scale_down") %>%
  landscape() %>%
  footnote(general = "C2 - corn in the 2-year rotation, C3 - corn in the 3-year rotation, C4 - corn in the 4-year rotation, S2 - soybean in the 2-year rotation, S3 - soybean in the 3-year rotation, S4 - soybean in the 4-year rotation, O3 - oat in the 3-year rotation, O4 - oat in the 4-year rotation, and A4 - alfalfa in the 4-year rotation.", threeparttable = TRUE, footnote_as_chunk = TRUE)

```


