---
output: 
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
    number_sections: false
    extra_dependencies: ["float"]
bibliography: ecol.bib
---

Diagnosis plots for all the statistical models used in "Impact of cropping system diversification on vegetative and reproductive characteristics of waterhemp (Amaranthus tuberculatus)" with `ggResidpanel` [@goodeGgResidpanelPanelsInteractive2019].


```{r, include=FALSE}
library(emmeans)
library(here)
library(tidyverse)
library(lme4)
library(magrittr)
library(ggResidpanel)
library(patchwork)
library(magick)
knitr::opts_chunk$set(message = FALSE, 
                      error = FALSE, 
                      warning = FALSE, 
                      fig.width = 8, #change number specs to balance plot size and resolution
                      fig.height = 8,
                      fig.fullwidth = T,
                     # fig.path = here("4 Graphs/"), 
                      dev=c("png"),
                      fig.pos = "H", 
                      table.placement = "H")

# Always log-transform possitive data: https://statmodeling.stat.columbia.edu/2019/08/21/you-should-usually-log-transform-your-positive-data/
```

#### Crop yield at the experiment site {-}

Crop identity in these yield models represented rotation system (2-year, 3-year or 4-year) for corn and soybean or 3-year and 4-year for oat.  

```{r, include=FALSE}  
### Experiment site crop yields
corn <- read_csv("../2-Data/Clean/corn_1720_clean.csv")
corn  %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

soy <-  read_csv("../2-Data/Clean/soybean_1720_clean.csv")
soy  %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

oat <- read_csv("../2-Data/Clean/oat_g_1720_clean.csv")
oat  %<>%  mutate_at(c("Block","Crop_ID", "Year"),funs(factor(.)))

```


```{r corn-mod, echo = TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on corn yield over four years with four blocks of replication."}
## Did crop identity and corn weed management affect corn yield?

corn.lmer <- lmer(log(Standardized_yield_MgpHa) ~ Block + 
                    Crop_ID*Corn_weed_management + 
                    (1|Year) +
                    (1|Year:Block) + 
                    (1|Year:Crop_ID) + 
                    (1|Year:Corn_weed_management) + 
                    (1|Year:Crop_ID:Corn_weed_management)  +
                    (1|Block:Year:Crop_ID) ,
  data=corn) #corn was harvested on hafl-plot basis

resid_panel(corn.lmer, "all")
```

```{r soy-mod, echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on soybean yield over four years with four blocks of replication."}
## Did crop identity and corn weed management affect soybean yield?

soy.lmer <- lmer(log(Standardized_yield_MgpHa) ~ Block + 
                   Crop_ID*Corn_weed_management + 
                   (1|Year) + 
                   (1|Year:Block) + 
                   (1|Year:Crop_ID) + 
                   (1|Year:Corn_weed_management) +
                   (1|Year:Crop_ID:Corn_weed_management)  + 
                   (1|Block:Year:Crop_ID),
  data=soy) #soybean was harvested on hafl-plot basis

resid_panel(soy.lmer, "all")
```


```{r oat-mod, echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity on oat yield over four years with four blocks of replication."}
## Did crop identity affect oat yield?
# crop identity represented rotation system (3-year or 4-year)

oat.lmer <- lmer(log(Standardized_yield_MgpHa) ~ Block + 
                   Crop_ID + 
                   (1|Year) +
                   (1|Year:Block) +  
                   (1|Year:Crop_ID) +
                   (1|Block:Year:Crop_ID),
  data=oat) #oat was harvested in whol-plot basis 

resid_panel(oat.lmer, "all")
```


#### Community ecological indices {-}  

Crop identities in these ecological indices models were the combinations of the crop species names' one-letter abbreviation and the rotation to which the crop belonged.

```{r, include=FALSE}
dens_ind_1720 <- read.csv(here("2-Data/Clean/pldens_indices_1720_clean.csv"))

#convert all id columns to factor 
dens_ind_1720 %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

dens_ind_1720$Crop_ID <- factor(dens_ind_1720$Crop_ID,
                                levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

dens_ind_1720$Crop <- factor(dens_ind_1720$Crop,
                             levels = c("corn", "soybean", "oat", "alfalfa"))

#dens_indices_1720 <- dens_ind_1720[,-c(1,2,3,8,13,14,16,17)] 

#dens_ind_1720 <- dens_ind_1720[rowSums(dens_ind_1720[, 12:15]) > 0, ]
```


```{r dens-div-mod, echo = TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community density diversity index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community diversity index?

dens_diversity.lmer1 <- lmer(log(Diversity + 1) ~ Block + 
                               Crop_ID*Corn_weed_management + 
                               (1|Year) + (1|Year:Block) + 
                               (1|Year:Crop_ID) + 
                               (1|Year:Corn_weed_management) +
                               (1|Year:Crop_ID:Corn_weed_management)  +
                               (1|Block:Year:Crop_ID) , 
                   data = dens_ind_1720,
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 
# summary(dens_diversity.lmer1)$sigma #0.27
resid_panel(dens_diversity.lmer1, "all")
```

```{r dens-even-mod, echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community density evenness index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community evenness index?

#min(dens_ind_1720$Evenness[dens_ind_1720$Evenness > 0])  #0.016156463
dens_even.lmer4 <- lmer(log(Evenness + 0.016156463) ~  Block + 
                          Crop_ID * Corn_weed_management + 
                          (1|Year) + 
                          (1|Year:Block) + 
                          (1|Year:Crop_ID) +
                          (1|Year:Corn_weed_management) +
                          (1|Year:Crop_ID:Corn_weed_management)  +
                          (1|Block:Year:Crop_ID) , 
                   data = dens_ind_1720,
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 
#summary(dens_even.lmer4)$sigma # 0.68 # second best sigma, better than arcsin sqrt transform and more spread out points

resid_panel(dens_even.lmer4, "all")
```

```{r dens-rich-mod, echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community density richness index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community density richness index?  

dens_rich.lmer2 <- lmer(log(Richness+1) ~ Block + 
                          Crop_ID * Corn_weed_management + 
                          (1|Year) +
                          (1|Year:Block) + 
                          (1|Year:Crop_ID) +
                         (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management) + 
                         (1|Block:Year:Crop_ID) , 
                   data = dens_ind_1720, 
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 
# summary(dens_rich.lmer2)$sigma  #0.288
resid_panel(dens_rich.lmer2, "all" )
```


```{r, include=FALSE}
biom_ind_1720 <- read_csv("../2-Data/Clean/biom_indices_1720_clean.csv")

#convert all id columns to factor 
biom_ind_1720  %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

biom_ind_1720$Crop_ID <- factor(biom_ind_1720$Crop_ID, levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

# zero variance solutions in glmer and lmer models: https://rpubs.com/bbolker/6226
# Block is fixed
```

```{r biom-div-mod, echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community aboveground mass diversity index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community biomass diversity index?

# min(biom_ind_1720$Diversity[biom_ind_1720$Diversity > 0])
biom_diversity.lmer1 <- lmer(log(Diversity + 1 ) ~  Block + 
                               Crop_ID * Corn_weed_management + 
                               (1|Year) + 
                               (1|Block:Year) + 
                               (1|Year:Crop_ID) + 
                               (1|Year:Corn_weed_management) + 
                               (1|Year:Crop_ID:Corn_weed_management)  + 
                               (1|Block:Year:Crop_ID)  , 
                   data = biom_ind_1720,
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 
# summary(biom_diversity.lmer1)$sigma #0.25

resid_panel(biom_diversity.lmer1, "all")
```
 
```{r biom-even-mod,  echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community aboveground mass evenness index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community aboveground mass evenness index?
min(biom_ind_1720$Evenness[biom_ind_1720$Evenness > 0]) 
biom_even.lmer4 <- lmer(log(Evenness + 0.015101721 ) ~ Block + 
                          Crop_ID * Corn_weed_management + 
                        (1|Year) + (1|Year:Block) + 
                       (1|Year:Crop_ID) + (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  + 
                         (1|Block:Year:Crop_ID) , 
                   data = biom_ind_1720,
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 
#summary(biom_even.lmer4)$sigma # 0.72 # second best sigma, points more spread-out 

resid_panel(biom_even.lmer4, "all")
```

```{r biom-rich-mod,  echo=TRUE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community aboveground mass richness index over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community aboveground mass richness index?

# min( biom_ind_1720$Richness[ biom_ind_1720$Richness > 0])
biom_rich.lmer2 <- lmer(log(Richness + 1) ~  Block + 
                          Crop_ID * Corn_weed_management + 
                        (1|Year)  +
                          (1|Year:Block) + 
                       (1|Year:Crop_ID) +
                         (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  +
                         (1|Block:Year:Crop_ID) , 
                   data = biom_ind_1720, 
                   control=lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4))) 

#summary(biom_rich.lmer2)$sigma  #0.2935

resid_panel(biom_rich.lmer2 , "all")
```



#### Total weed community density and aboveground mass {-}

```{r, include=FALSE}
# Read in community total biomass 
total_biom_17_20 <- read_csv("../2-Data/Clean/biom_1720_clean.csv")

total_biom <- total_biom_17_20[,-c(9:42)] #keep total abovegorund mass only

total_biom <- total_biom   %<>%
  mutate_at(c("Block","Crop_ID", "Corn_weed_management",  "Year"),
            funs(factor(.)))

total_biom$Crop_ID <- factor(total_biom$Crop_ID,
                             levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

# Read in community total density
total_dens_17_20 <- read_csv(here("2-Data/Clean/pldens_1720_clean.csv"))
total_dens <- total_dens_17_20[,-c(9:42)] #keep total density only
total_dens <- total_dens   %<>%
  mutate_at(c("Block","Crop_ID", "Corn_weed_management", "Year"),
            funs(factor(.)))

total_dens$Crop_ID <- factor(total_dens$Crop_ID,
                             levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))
```

```{r all-biom-mod, echo=FALSE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community aboveground mass over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community aboveground mass?

# min( total_biom$Total[ total_biom$Total > 0]) #0.00108
comm_biom_lmer <- lmer(log(Total + 0.00108)  ~ Block + 
                        Crop_ID * Corn_weed_management + 
                        (1|Year)  +
                        (1|Year:Block) + 
                       (1|Year:Crop_ID) + 
                        (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  +
                        (1|Block:Year:Crop_ID) ,
                   data = total_biom) 


resid_panel(comm_biom_lmer, "all")
```

```{r all-dens-mod, echo=FALSE, fig.cap="Diagnosis plot for the effect of crop identity and corn weed management on weed community density over four years with four blocks of replication."}
## Did crop identity and corn weed management affect weed community density?

# min( total_dens$Total[ total_dens$Total > 0]) #0.054
comm_dens_lmer <- lmer(log(Total + 0.054)  ~ Block +
                        Crop_ID * Corn_weed_management + 
                        (1|Year)  + 
                        (1|Year:Block) + 
                       (1|Year:Crop_ID) + 
                        (1|Year:Corn_weed_management) + 
                       (1|Year:Crop_ID:Corn_weed_management)  +
                        (1|Block:Year:Crop_ID) ,
                   data = total_dens) 

resid_panel(comm_dens_lmer, "all")
```

#### Top seven species individual density {-}


```{r, include=FALSE}
# Read in individual weed species density 
dens_1720 <- read.csv(here("2-Data/Clean/pldens_1720_clean.csv"))
dens_1720_clean <- dens_1720 %>%
  replace(is.na(.), 0) %>%
  mutate(zOTHERS = Total - (AMATA + CHEAL + DIGSA + ECHCG + SETFA + SETLU + TAROF)) %>% #keep top seven species and sum all other species in to zOTHERS
  dplyr::select(Crop, Crop_ID, Block, Year, Corn_weed_management,
    AMATA , CHEAL , DIGSA , ECHCG , SETFA , SETLU , TAROF, zOTHERS) %>% 
  mutate_at(13, funs(round(., 8))) # zOTHERS make the group listed last in ggplot2


biom_1720 <- read.csv(here("2-Data/Clean/biom_1720_clean.csv"))
biom_1720_clean <- biom_1720 %>% 
  replace(is.na(.), 0)%>%
   mutate(zOTHERS = Total - (AMATA + CHEAL + DIGSA + ECHCG + SETFA + SETLU + TAROF)) %>% #keep top seven species and sum all other species in to zOTHERS
  dplyr::select(Crop, Crop_ID, Block, Year, Corn_weed_management,
                AMATA , CHEAL , DIGSA , ECHCG , SETFA , SETLU , TAROF, zOTHERS) %>%
  mutate_at(13, funs(round(., 8))) # zOTHERS make the group listed last in ggplot2


# remove rows that sum to zeros https://www.tutorialspoint.com/how-to-remove-rows-that-contains-all-zeros-in-an-r-data-frame
#dens_1720_clean <- dens_1720[rowSums(dens_1720[,13:46])>0,] 

dens_1720_clean$Crop <- factor(dens_1720_clean$Crop, levels = c("corn", "soybean", "oat", "alfalfa"))
dens_1720_clean$Crop_ID <- factor(dens_1720_clean$Crop_ID, levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

#biom_1720_clean <- biom_1720[rowSums(biom_1720[,13:46])>0,] 
biom_1720_clean$Crop <- factor(biom_1720_clean$Crop, levels = c("corn", "soybean", "oat", "alfalfa"))

biom_1720_clean$Crop_ID <- factor(biom_1720_clean$Crop_ID, levels = c("C2", "S2", "C3", "S3", "O3", "C4", "S4", "O4", "A4"))

dens_1720_clean %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

biom_1720_clean %<>%  mutate_at(c("Block","Crop_ID", "Year","Corn_weed_management"),funs(factor(.)))

# Find minimum non-zero in each matrix, or in the Prop column of the long-format dframe  
```


```{r, echo=TRUE}

##### Fit all species at once with split-fit-combine syntax 
# https://stat585-at-isu.github.io/materials-2019/04_functional-programming/02_purrr.html#23


# convert wide to long format 

dens_1720_long <- dens_1720_clean %>% 
  pivot_longer(!c(Crop:Corn_weed_management), 
               names_to = "Species", values_to =  "Density")

#singularity because of too many zeros in many species columns 


# min(dens_1720_long$Density[dens_1720_long$Density>0]) #0.05396072

dens_result_t <- dens_1720_long %>%
  group_by(Species) %>% 
  nest() %>%
  mutate(models=map(data,~lmer(log(Density + 0.05396072) ~ Block + Crop_ID +
                                 Corn_weed_management  +
                                 Crop_ID:Corn_weed_management + 
                                 (1|Year) + (1|Year:Block) + 
                                 (1|Year:Crop_ID) + 
                                 (1|Year:Corn_weed_management) + 
                                 (1|Year:Crop_ID:Corn_weed_management)  +
                                 (1|Block:Year:Crop_ID) ,data =.x))) %>%
  ungroup %>%
  mutate(diag_plots = map(models, resid_panel, "all"),
    jts = map(models, joint_tests),
         across(diag_plots:jts, setNames,  .$Species))
```


```{r sp-dens-diag, echo=FALSE, fig.cap= "Diagnosis plot for the effect of crop identity and corn weed management on the aboveground mass of (A) - AMATA, (B) - CHEAL, (C) - DIGSA, (D) - ECHCG, (E) - SETFA, (F) - SETLU, and (G) - TAROF", fig.height=22, fig.width=22}

list(dens_result_t$diag_plots$AMATA,
dens_result_t$diag_plots$CHEAL,
dens_result_t$diag_plots$DIGSA,
dens_result_t$diag_plots$ECHCG,
dens_result_t$diag_plots$SETFA,
dens_result_t$diag_plots$SETLU,
dens_result_t$diag_plots$TAROF) %>%
  wrap_plots() +
  plot_annotation(tag_levels = "A")
```

#### Top seven species individual aboveground mass {-}

```{r, echo=TRUE}

biom_1720_long <- biom_1720_clean %>% 
  pivot_longer(!c(Crop:Corn_weed_management), 
               names_to = "Species", values_to =  "Biomass")

# min(biom_1720_long$Biomass[biom_1720_long$Biomass>0]) # 0.0005396072

biom_result_t <- biom_1720_long %>%
  group_by(Species) %>% 
  nest() %>%
  mutate(models=map(data,~lmer(log(Biomass + 0.0005396072) ~ Block + 
                                 Crop_ID +  Corn_weed_management +
                                 Crop_ID:Corn_weed_management +
                                 (1|Year) + (1|Year:Block) + 
                                 (1|Year:Crop_ID) + 
                                 (1|Year:Corn_weed_management) + 
                                 (1|Year:Crop_ID:Corn_weed_management)  +
                                 (1|Block:Year:Crop_ID),data =.x))) %>%
  ungroup %>%
  mutate(jts = map(models, joint_tests),
         diag_plots = map(models, resid_panel,"all"), 
across(jts:diag_plots, setNames,  .$Species))

```

```{r sp-biom-diag, echo=FALSE, fig.cap= "Diagnosis plot for the effect of crop identity and corn weed management on the density of (A) - AMATA, (B) - CHEAL, (C) - DIGSA, (D) - ECHCG, (E) - SETFA, (F) - SETLU, and (G) - TAROF", fig.height=22, fig.width=22}

list(biom_result_t$diag_plots$AMATA,
biom_result_t$diag_plots$CHEAL,
biom_result_t$diag_plots$DIGSA,
biom_result_t$diag_plots$ECHCG,
biom_result_t$diag_plots$SETFA,
biom_result_t$diag_plots$SETLU,
biom_result_t$diag_plots$TAROF) %>%
  wrap_plots() +
  plot_annotation(tag_levels = "A")
```


