---
title: "Weed community composition in simple and more diverse cropping systems"
header-includes:
   - \usepackage{lineno}
   - \linenumbers
output:
  bookdown::pdf_document2: 
    keep_tex: yes
    toc: false
    number_sections: false
    extra_dependencies: ["float"]
    latex_engine: lualatex
    citation_package: natbib
bibliography: ecol.bib
natbiboptions: round
biblatexoptions:
  - sortcites = true
  - sorting = nyt
  - backend = biber
  - maxcitenames = 2
csl: apa-no-ampersand.csl
---

# Abstract {-}

```{r, child = "Abstract.Rmd"}
```

# Introduction {-}

```{r, child = "Introduction.Rmd"}
```


# Materials and Methods {-}


```{r, child = "MandM.Rmd"}
```

# Results {-}  

A lack of any obvious bias in plots of residuals versus predicted values suggested that the analysis models fit the data well. Diagnosis plots made with `ggResidpanel` version 0.3.0 [@goodeGgResidpanelPanelsInteractive2019] are available in the Data Repository.  


```{r, child = "../5-Analysis/Crop_yields.Rmd"}
```


```{r, child = "../5-Analysis/Community.Rmd"}
```


```{r, child = "../5-Analysis/Individual-sp-difference.Rmd"}
```



# Discussion {-}

```{r, child = "Discussion.Rmd"}
```


# Acknowledgements {-}

The authors thank Matt Woods, Mike Fiscus, and the Iowa State University's Agronomy Research Farm crew for field management; Wendy Borja-Diaz, Lydia English, Jessica Juarez-Morales, Samantha Kanselaar, Jessica Nelson, Elizabeth Oys, Ana Poznanski, Andrew Riehl, Angela Soto-Saenz, Mickala Stallman, David Weisberger, and Wyatt Westfall for field and laboratory assistance; Katherine Goode, Audrey McCombs, Philip Dixon and ISU's statistical consulting group for data analysis assistance; Russ Lenth and other Stackoverflow community members for answering HTXN's coding questions; Micheal Owen for reviewing the manuscript; Overleaf staff for \LaTeX\ assistance in compiling the manuscript; and two anonymous reviewers for their constructive feedback.


# References {-}
